Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework

### 2015-03-24 tex/manuscript.tex
Major revisions in prep for GEB submission

%\usepackage[margin=1in, left=1.2in]{geometry}

@@

\usepackage{caption}
\captionsetup{labelformat=simple, labelsep=none}

@@

%% graphics packages
\usepackage{graphicx}

% Tikz libraries for building the diagram
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc, shapes}
\usetikzlibrary{shapes.geometric,shapes.arrows,decorations.pathmorphing}
\usetikzlibrary{matrix,chains,scopes,positioning,arrows,fit}

@@

%% ----------------------------------
%
%     Packages to be used for drafts and editing
%     Remove from the final manuscript
%
%% ----------------------------------
\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}

@@

	\textbf{Abstract word count: } 258

	\textbf{Main text word count: } 6197 %WORD_COUNT %Target 5000

@@

Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Guisan2013}.
Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}. 
Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, evolution \citep{Jay2012}, hybrid zone dynamics \citep{Engler2013}, and forecasting the impacts of climate change on species distributions \citep{Thuiller2011, Blois2013, Thuiller2014}.

Despite the recognized potential of these models, it can be difficult to produce predictions with acceptable levels of precision and bias \citep{Guisan2013}.
For mechanistic models, two important constraints can present difficulties in reducing uncertainty and bias: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
In recent decades, however, modeling techniques have proliferated to take advantage of increased availability of data. 
A growing body of theory, reflecting the diversity of processes generating species ranges, also has contributed to model diversification \citep{Boulangeat2012}.
Of these model types, fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, perform poorly when applied at the scale of species ranges.
For instance, biotic interactions are usually not modeled mechanistically at the scale of species distributions because they are poorly known or have not been recorded, despite being considered a key determinant of range limits \citep{Holt2005, Pigot2013}. 
In contrast to mechanistic models, more correlative approaches that statistically relate species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}. 
Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
%% ----------------------------------
%% ----------------------------------
%% ----------------------------------

\subsection*{Toward an integrated approach for modeling ranges}
\mt{this transition is fairly abrupt. it would be nice to have more of an intro explaining the motivation. perhaps it comes later in the intro; if so, it should be moved here}
\mt{a structure to consider:\\
1. intro sentence (gee it would be great to combine data sources) \\
2. "Some researchers... hierarchical nature..." (but make it less awful) \\
3. Hybrid models (and rewrite -- it is a bit redundant) \\
4. Ensemble forecasting
5. Our approach
-- could break this into different paragraphs
}
To include knowledge of species from multiple sources and scales into distribution models, we present an application of hierarchical Bayesian methods that uses outputs from multiple model outputs to inform the results of the final model.
Our approach is an alternative to other approaches to multimodel inference that offers improved flexibility in incorporating data and theory from multiple scales and provides transparent estimation of uncertainty.
By integrating all available knowledge while producing a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Pearson2003, Guisan2005, Araujo2006, Quillet2010}.
\mt{Blech. This transition is also terrible}
Some researchers have explicitly reproduced the hierarchical nature of ecological systems using hierarchical models \citep[e.g.][]{Royale2008, Boulangeat2012, Catterall2012, Strigul2012, Soranno2014}. 
Recent literature has also focused on the development of hybrid models, which allow for flexible combinations between mechanistic and phenomenological sub-models \citep{Gallien2010, Thuiller2013}. 
For instance, correlative models are used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while mechanistic approaches can include space and time dynamics and therefore account for dispersal processes \citep{Kearney2008}. 
This complementarity has been used to merge the two model types into hybrid models \citep[e.g.][]{Keith2008, Anderson2009, Smolik2010, Dormann2012, Boulangeat2014}. 
However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, and the sources and magnitudes of uncertainties are difficult to quantify. 
Moreover, hybrid approaches are not suitable for merging models that are based on the same processes but with different underlying hypotheses.

A commonly-employed alternative to hybrid models is the direct combination of predictions.
Predictions from models operating at the same spatio-temporal scales and having compatible parameterizations can be combined using multi-model inference \citep[e.g., model averaging, ensemble forecasting;][]{Thuiller2004, Araujo2007}. 
However, the applicability of such techniques is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.
A more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying the parameters that contribute most to variance in the model predictions \citep{McMahon2011}. 
Clearly, we require an approach that can both (1) flexibly incorporate multiple modes of inference (i.e., mechanistic, correlative models), (2) encompasses data from multiple sources and scales, and (3) explicitly includes uncertainty from all sources \citep{Levin1992, Peters2004, Beck2012, Thuiller2013}.

Here, we present a framework for modeling species range dynamics that aims to account for all the information available on a focal species, from a variety of sources and scales.
Contrary to hybrid and hierarchical models, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for as much flexibility as possible regarding the type of information included. 
We use the power and generality of a hierarchical Bayesian framework, which allows us to include multiple data sources and modes of inference \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}. 
Another advantage of Bayesian methods is that uncertainty in model outputs is intuitive to interpret and reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}. 
We illustrate our approach with two examples.
We begin with a hypothetical example, using simulated data, that presents a simplified application of our framework to multiple sources of information from different scales (Example 1).
In a second example, we apply the framework to combine presence-absence information with phenological data to improve uncertainty estimates and reduce bias when predicting changes to the range of sugar maple (\emph{Acer saccharum} Marsh.), a widespread foundation tree species from eastern North America, in response to climate change.
Both examples are presented with complete code and data (in Supporting Information).
Finally, we provide a more formal mathematical presentation in Appendix S1.

%% ----------------------------------
%
%     MODELING FRAMEWORK
%
%% ----------------------------------

\section*{Integrated modeling framework}

The key idea of our approach is formulate a metamodel that integrates data at the same ecological scale as the desired predictions.
The parameters of this model can be constrained using the output of one or more sub-models that can operate at different scales than the metamodel.
As an example, the metamodel could take the form of a simple correlative species distribution model (SDM) operating at very large spatial scales, with one or more fine-scaled process-based sub-models.
Ideally, each sub-model incorporates different data types and hypotheses that specify underlying ecological mechanisms contributing to the model output.
The output of sub-models should provide additional predictions at a comparable ecological scale at which the metamodel operates, although it is not necessary for the sub-models to actually operate at the same scale.
For sub-models operating at a substantially different scale than the metamodel, scaling functions will be required; however, these functions can be simple while still contributing useful information to the metamodel (see Example 1). 
All predictions may therefore be integrated in a single Bayesian framework when evaluating the parameters of the metamodel (see Appendix S1, Fig. \ref{fig:diagram}).
We present and apply the framework here with two examples, and follow with a more formal presentation in Appendix S1.
\mt{There is still some repetition here -- two refs to Appendix S1 in 2 sentences}

%% ----------------------------------
%
%     EXAMPLE 1
%
%% ----------------------------------
\subsection*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}

In this hypothetical example, we build a metamodel relating the distribution (i.e. occurrence probability) of an annual plant to coarse-scale climate with complementary information originating from a fine-scale experiment manipulating the precipitation regime.
The metamodel attempts to capture the realized distribution of a species; as a correlative model, it implicitly captures the major physiological constraints and ecological processes constraining the distribution of the target species. 
However, for the purposes of forecasting, we would like to disentangle the fundamental response of a species to environmental variation from other processes in order to map the climatic envelope of where a species may be found in a natural setting. 
Prior information of the physiological constraints affecting species distribution is sometimes available, but usually too incomplete to perform a reasonable comparison with the realized distribution.
For instance, as in this example, a species distribution might be constrained by both temperature and precipitation, but experiments were conducted only over a precipitation gradient at one temperature. 
We apply our framework here to such heterogeneous sources of information to reduce the bias in parameter estimation and more adequately represent uncertainty (see Appendix S2 in Supporting Information for procedural details and scripts for executing the model).

We consider data collected from a species' historical distribution, where the goal is to predict the distribution following a substantial reduction in precipitation. 
We will consider \(X_M\) (see Appendix S1, Fig. \ref{fig:diagram}) to be a vector of $n$ observations that takes the value of one to indicate presence and zero for absence:
\mt{will need to edit this, since \(X_M\) is no longer defined clearly in the text (has been moved to appendix S1)}
\begin{equation}
X_M = \{x_{M,1}, x_{M,2}, \ldots, x_{M,n}\}
\end{equation}
We assume the data are relatively high-quality, providing coverage over a wide region and covering various climatic conditions (Fig. \ref{fig:ex1_sampling}a). 
The model is evaluated by relating these observations to environmental data \((D_M)\), which, for the sake of the example, consists of mean annual temperature $T_M$ and annual precipitation $P_M$. 
For simplicity, we formulate the metamodel \((\theta_M)\) as a simple logistic regression with a second order effect of temperature and precipitation. 
This naive model (i.e., the metamodel with no constraints from sub-models) predicts  occurrence probability (\(\psi_N\)) as a function of the environment:
%-----
\begin{equation}
\begin{aligned}
	\psi_N &= f\left(\theta_M, D_M \right) \\
	&= p \left (X_M = 1 \mid \theta_M, T_M, P_M \right) \\
	&=\text{logit}^{-1}\left( \mathbf{\Theta_M} \mathbf{D_M} \right)
\end{aligned}
\end{equation}
%-----
where \(\mathbf{\Theta_M}\) is the parameter vector of the model \(\theta_M\), \(\mathbf{D_M} \) is the covariate matrix (i.e., \(T_M, P_M\), with the first column taken to be unity to allow an intercept to be fit), and \(\text{logit}^{-1}\) is the inverse of the logit function.
We can fit this model in a Bayesian framework to allow for easy integration of models (as we will show later).
In this context, the goal of modeling is to estimate the probability distribution of \(\theta_M\), the model parameters, given the observed data \((X_M, T_M, P_M)\), which is given by the proportional form of Bayes's Theorem \citep[for readers unfamiliar with general concepts in Bayesian inference, a concise introduction can be found in ][]{Link2010}:
%-----
\begin{equation}
\label{eq:ex1_bayes}
	p\left (\theta_M \mid X_M,T_M,P_M \right ) \propto 
	p \left(X_M \mid \theta_M, T_M, P_M \right)
	p \left(\theta_M \right)
\end{equation}
%-----
where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.

Thus far, we have considered only a single source of information to fit this model, and therefore the prior distribution \(p\left(\theta_M \right)\) from Eq. \ref{eq:ex1_bayes} is uninformative.
As a secondary source of information, we will consider an experiment relating the population growth rate of the plant to manipulations to the precipitation regime, with results (but no raw data) available from the literature (Fig. \ref{fig:ex1_sampling}b). 
Furthermore, no information is available regarding the temperature regime for the experiment.
Transplant experiments that evaluate performance beyond the range of a species are common, and represent a plausible scenario for model integration \citep{Hargreaves2014}.
According to niche theory \citep{Holt2009}, the fundamental niche corresponds to the set of environmental conditions where the per capita intrinsic growth rate $r$ is positive.
This concept gives us a reasonable model to fit the scaling function $g$ for our sub-model (Fig. \ref{fig:diagram}).
If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed, then for observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
\begin{equation}
	\psi_{S,i} = \int_0^\infty N \left(X_{S,i}, \sigma_{S,i} \right)
\end{equation}
where \(N\) is the Normal density function.
We can then estimate the posterior distribution for the sub-model by fitting the relationship between \(\psi_S\) and precipitation \( \left( P_S \right) \) using Bayesian beta regression \citep{Ferrari2004}:
%-----
\begin{equation}
\label{eq:ex1_thetas}
	p\left (\theta_S \mid \psi_S,P_S \right ) \propto 
	p \left(\psi_S \mid \theta_S, P_S \right)
	p \left(\theta_S \right) \\
\end{equation}
%-----

Although the two datasets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation). 
This scaling is quite simplistic, and would never be used to predict a species' range using this mechanistic sub-model in isolation.
Despite the strong theoretical foundation (i.e., niche theory) linking the observations to presence-absence at large scales, the upscaled sub-model (i.e., the transformation of \(X_S\) and \(\sigma_S\) to \(\psi_S\)) is built on the strong hypothesis that the species will be present when $r>0$, neglecting other population dynamics constraints such as Allee effects and metapopulation dynamics. 
The fundamental niche is also incomplete, as we lack environmental variables such as temperature, as well as all other ecological processes responsible for the realized distribution.
As such, it would be unwise to expect predictions from this model alone to resemble the actual distribution of the species; as a mechanistic model, it is simply too incomplete to predict distribution.
As we will show, however, the information from this sub-model, when applied as a constraint on the metamodel, can result in improved predictions that incorporate the information within each model.

We accomplish model integration by treating the posterior of \(\theta_S\) as prior information about some of the parameters of \(\theta_M\) (i.e., parameters related to precipitation), expanding Equation \ref{eq:ex1_bayes} to incorporate the new information from the sub-model:
%-----
\begin{equation}
	\label{eq:ex1_integrated}
	\overbrace{p(\theta_M \mid X_M, T_M, P_M, \theta_S, \psi_S, P_S)}^\text{integrated posterior}
	\propto
	\overbrace{p\left (\psi_S \mid \theta_S,P_S \right )}^{\substack{\text{new information} \\ \text{from sub-model}}}
	\overbrace{p \left(X_M \mid \theta_M, T_M, P_M \right) p \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
	\overbrace{p \left(\theta_S \right)}^{\substack{\text{prior for} \\ \text{sub-model}}}	
\end{equation}
%-----
As before, the metamodel \(\theta_M\) can be used to predict the species distribution \((\psi_I)\).
However, these predictions will not reflect only the presence-absence samples in \(X_M\), but will also reflect the information from \(\theta_S\), including all of the data sources used to produce this sub-model.
In other words, with this evaluation, the set of parameters $\theta_M$ is most likely when it agrees with both the original data and predictions from the sub-model. 
Note that in this particular example, because the sub-model is only contingent on precipitation, there is no integration on temperature. 
Finally, we note the presence of marginal distributions for both models (i.e., \(p(\theta_M)\) and \(p(\theta_S)\)).
These can be informative (e.g., incorporating further prior information or the predictions of additional sub-models), semi-informative (e.g., to provide greater weight to more informative models), or uninformative.
For purposes of this example, we used these terms to reduce the precision of the second model to reflect the uncertainty of the modeling process (due to the lack of additional mechanisms in the second model, as mentioned above).

We performed this procedure with our hypothetical datasets.
When comparing the three models (naive metamodel, mechanistic sub-model, and integrated metamodel), we observed extreme uncertainty in the first model when projecting beyond the range of the original data (Figs. \ref{fig:ex1_precip}a, \ref{fig:ex1_map}a, \ref{fig:ex1_map}b)
Unsurprisingly given the quality of the data, the sub-model was highly precise with respect to precipitation and not useful for temperature, thus providing a fairly strong constraint when producing the integrated model (Fig \ref{fig:ex1_precip}b).
The result was an integrated prediction that reflected the shapes of both models and showed considerably reduced uncertainty (Fig. \ref{fig:ex1_precip}).
At the scale of the metamodel, considering both temperature and precipitation, we observed similar results, with reduced uncertainty in the predictions over the domain not covered by the presence-absence data (Fig. \ref{fig:ex1_map}).

Once the technical challenges of evaluating the parameters are overcome, integration has several advantages over both the naive parameterization of the metamodel and the mechanistic sub-model. 
In the situation where both models strongly agree on the relationship between water availability and occurrence probability, then we should expect significantly reduced uncertainty in that domain (Figs. \ref{fig:ex1_precip}, \ref{fig:ex1_map}).
Because the metamodel includes both temperature and precipitation, any reduction in the uncertainty with respect to precipitation might also influence the evaluation of the relationship with temperature (reducing bias in the case where there is a correlation between T and P in the data). 
However, uncertainty could increase in the case of model disagreement.
For instance, the relationship between precipitation and presence in the naive metamodel could be caused by a spurious relationship (e.g. spatial autocorrelation caused by historical contingencies), leading to a false confidence in the species distribution over precipitation gradients. 
The information from the sub-model will thus bring back the right amount of uncertainty in the relationship.
Moreover, convergence in model predictions in overlapping domains (e.g., near the peak in the probability of presence in Fig. \ref{fig:ex1_precip}) can aid in detecting crucial ecological processes.

@@

%% ----------------------------------
%
%     FIGURES
%
%% ----------------------------------

%TC:break figures
\newpage
\section*{Figure Legends}
\begin{flushleft}

>>>>

\usepackage[margin=1in]{geometry}

@@

@@

%% graphics packages
\usepackage{graphicx}
%\usepackage[nomarkers,nolists,heads]{endfloat}
\usepackage{caption}
\captionsetup{labelformat=simple}
\usepackage{booktabs}
\usepackage{multirow}

@@

%% ----------------------------------
%
%     Packages to be used for drafts and editing
%     Remove from the final manuscript
%
%% ----------------------------------
%\usepackage[top=1in, left=1in, bottom=1in, right=2.5in]{geometry}
%\usepackage[textsize=tiny, backgroundcolor=white, textwidth=2.2in, colorinlistoftodos=true]{todonotes} % for margin notes using \todo{}
%\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}

@@

	\textbf{Abstract word count: } 258 %ABS_COUNT

	\textbf{Main text word count: } 4641 %WORD_COUNT %Target 5000

@@

Models of species range limits have wide applications and can play a large role in conservation biology, where they can be used as decision-making tools in biodiversity management \citep{Guisan2013}.
Due to large temporal and spatial scales as well as the complex and nonlinear nature of ecosystem dynamics, it is often impossible to construct experiments that can adequately explore the ecological processes generating species range limits \citep{Wu1995, Levin1998}. 
Hence, range models are essential tools that have been applied to a large number of ecological subfields, including biogeography \citep{Schurr2012}, invasion biology \citep{Catterall2012, Gallien2012}, hybrid zone dynamics \citep{Engler2013}, and forecasting the impacts of climate change on species distributions \citep{Blois2013, Thuiller2014}.

Despite the recognized potential of these models, it can be difficult to produce species distribution models (SDMs) with acceptable levels of precision and bias \citep{Guisan2013}.
For mechanistic models, two important constraints can present difficulties: (1) having the appropriate ecological theory needed to link data to modeling objectives, and (2) having sufficient data over a range of conditions to maintain coherence between the spatial and temporal scales of data and theory.
In recent decades, however, modeling techniques have proliferated to take advantage of increased data availability. 
A growing body of theory, reflecting the diversity of processes generating species ranges, also has contributed to model diversification \citep{Boulangeat2012}.
Of these model types, fine-scale mechanistic models may capture important ecological processes quite well, but due to their specificity, perform poorly when applied at the scale of species ranges.
For instance, biotic interactions are usually not modeled mechanistically at the scale of species distributions because they are poorly known or have not been recorded, despite being considered a key determinant of range limits \citep{Pigot2013}. 
In contrast to mechanistic models, more correlative approaches that statistically relate species occurrences to associated distributions of other variables have the advantage of indirectly accounting for underlying processes \citep{Guisan2000}.
However, their predictions rely on the constancy of the relationships between occurrences and explanatory variables in time and space, implying that the selected variables are related to the processes limiting species ranges and that their correlations are constant for calibration and projection ranges in space and time \citep{Dormann2007}. 
Extrapolating beyond the scope of the original data (e.g., predicting ranges based on future climate) is therefore problematic, because nonlinearities in responses to novel combinations of the explanatory variables cannot be accommodated in models that do not simulate the underlying processes.
%% ----------------------------------
%% ----------------------------------
%% ----------------------------------

Clearly, an approach is needed that unifies the strength of the diversity of modeling approaches and can incorporate multiple data sources and modeling paradigms.
Toward this end, we present here an application of hierarchical Bayesian methods that uses outputs from multiple model outputs to inform the results of the final model.
Techniques for multimodel inference have proliferated in recent years.
For example, hybrid models that allow for flexible combinations between mechanistic and phenomenological sub-models are commonly employed in SDMs \citep{Gallien2010, Thuiller2013, Boulangeat2014}.
Within the hybrid framework, a correlative model might be used to account for abiotic variables limiting species distributions \citep{Guisan2005}, while a more mechanistic approach could include space and time dynamics and therefore account for dispersal \citep{Smolik2010}.
However, the link between different sub-models is based on assumptions about the scaling of ecological processes that are poorly known and difficult to test \citep{Gallien2010}, and the sources and magnitudes of uncertainties are difficult to quantify. 
A commonly-employed alternative to hybrid models is the direct combination of predictions, where predictions from models operating at the same spatio-temporal scales can be combined \citep[e.g., model averaging, ensemble forecasting;][]{Araujo2007}. 
However, the applicability of such techniques is limited; for example, it is not currently possible to evaluate the effects of convergent predictions on the total uncertainty of the outcomes, despite its potential importance in a prediction context.

We propose an alternative to these other approaches to multimodel inference.
We use the power and generality of a hierarchical Bayesian framework, which provides a number of desirable advantages, including (1) the ability to incorporate multiple modes of inference (e.g., mechanistic, correlative models) \citep{VanOijen2005, Clark2006, Hobbs2011, Hartig2012}, (2) an easy mechanism to include multiple data sources at various scales \citep{Levin1992, Peters2004}, and (3) intuitive and comprehensive reporting of uncertainty in model predictions that  reflects uncertainty at all levels of organization \citep{Cressie2009, Hobbs2011}.
Contrary to hybrid and ensemble methods, the aim is not to link different sub-models into a single model, but to condition the predictions of a metamodel at the target scale (e.g., an entire species' range) with information from independent sub-models at a variety of spatial scales, allowing for as much flexibility as possible regarding the type of information included. 
By integrating all available knowledge while producing a single prediction, our approach potentially mitigates the limitations inherent in each individual model, contributing to more robust predictions \citep{Guisan2005, Araujo2006}.
Moreover, a more comprehensive understanding of uncertainty can guide biodiversity management and prioritize future data collection by identifying the parameters that contribute most to variance in the model predictions \citep{McMahon2011}.

We illustrate our approach with two examples.
We begin with a hypothetical example, using simulated data, where we define the framework and demonstrate its application to multiple sources of information from different scales.
In a second example, we apply the framework to combine presence-absence information with phenological data to improve uncertainty estimates and reduce bias when predicting changes to the range of sugar maple (\emph{Acer saccharum} Marsh.), a widespread foundation tree species from eastern North America, in response to climate change.
Both examples are presented with complete code and data (in Supporting Information).
Finally, we provide a more formal mathematical presentation in Appendix S1.

%% ----------------------------------
%
%     EXAMPLE 1
%
%% ----------------------------------
\section*{Example 1: Adding experimental evidence for the fundamental niche to a species distribution model}
The key idea of our approach is formulate a metamodel that integrates data at the same ecological scale as the desired predictions, and to constrain the parameters of this model using the output of one or more sub-models
In this hypothetical example, we build a metamodel relating the distribution (i.e. occurrence probability) of an annual plant to coarse-scale climate with complementary information originating from a fine-scale experiment manipulating the precipitation regime.
The metamodel attempts to capture the realized distribution of a species; as a correlative model, it implicitly captures the major physiological constraints and ecological processes constraining the distribution of the target species. 
However, for the purposes of forecasting, we would like to disentangle the fundamental response of a species to environmental variation from other processes in order to map the climatic envelope of where a species may be found in a natural setting.
Thus, we use additional information on the physiological constraints affecting species distribution.
Because these data are often collected at a finer scale than that of rangewide occurence data, we apply a simple scaling function, drawing on ecological theory, to compute the likelihood of a set of metamodel parameters given both the occurrence and the physiological data.
See Appendix S1 in Supporting Information for procedural details and scripts for executing the model.

%==================
%TC:ignore
\begin{figure}[t]
	\includegraphics{figs/ex1_sampling.pdf}
	\caption{Two simulated datasets used to illustrate the model integration framework.
	(a) Presences (circles) or absences (x's) of the species in ecological space, where precipitation ranges from 0--1.
	(b) Growth rate ($r$) as a function of manipulations to the precipitation regime (whiskers show $\pm$ 1 SE), with a larger range for precipitation (i.e., \(-\)1--1).
	The dashed line shows the threshold above which the species net growth rate is positive (implying presence).
	Axis scales for temperature and precipitation are arbitrary, but note the different scales on the horizontal axes.}
	\label{fig:ex1_sampling}
\end{figure}
%TC:endignore
%==================

We consider data collected from a species' historical distribution, where the goal is to predict the distribution following a substantial reduction in precipitation.
We assume a presence-absence dataset, which we term \(X_M\), where the subscripted \(M\) indicates that the data were collected at the same scale as the metamodel.
We simulated a relatively high-quality dataset that provides coverage of a variety of ecological conditions (Fig. \ref{fig:ex1_sampling}).
We desire to model the species' distribution as a function of temperature, \(T_M\), and precipitation, \(P_M\).
An initial version of the metamodel \((\theta_M)\) that has no constraints from other datasets will be referred to as the naive model.
This naive model uses a simple Bayesian logistic regression to estimate the naive probability of occurrence \((\psi_N)\) as a function of second order effects of temperature and precipitation:
%-----
\begin{equation}
\begin{aligned}
	\psi_N &= f\left(\theta_M, T_M, P_M \right) \\
	&= p \left (X_M = 1 \mid \theta_M, T_M, P_M \right) \\
	&=\text{logit}^{-1}\left( \theta_M D_M \right)
\end{aligned}
\end{equation}
%-----
where \(theta_M\) is the parameter vector of the model, \(D_M\) is the covariate matrix (i.e., \(T_M, P_M\), and quadratic terms, with the first column taken to be unity to allow an intercept to be fit), and \(\text{logit}^{-1}\) is the inverse of the logit function.
We estimate parameters using a Markov Chain Monte Carlo (MCMC) method using the proportional form of Bayes' theorem:
%-----
\begin{equation}
\label{eq:ex1_bayes}
	p\left (\theta_M \mid X_M,T_M,P_M \right ) \propto 
	p \left(X_M \mid \theta_M, T_M, P_M \right)
	p \left(\theta_M \right)
\end{equation}
%-----
where \(p\left(X_M \mid \theta_M, T_M, P_M \right)\) is often referred to as the \emph{likelihood} of the data (\(X_M\)) given the model (\(\theta_M\)), \(p\left(\theta_M \right)\) is often referred to as the \emph{prior distribution} of \(\theta_M\), and the goal of modeling is to estimate \(p\left (\theta_M \mid X_M,T_M,P_M \right )\), the \emph{posterior distribution} of \(\theta_M\), which gives the probability that \(\theta_M\) takes particular values, given the observed data.

Thus far, we have considered only a single source of information to fit this model, and therefore the prior distribution \(p\left(\theta_M \right)\) from Eq. \ref{eq:ex1_bayes} is uninformative.
As a secondary source of information, we will consider an experiment relating the population growth rate of the plant to manipulations to the precipitation regime, with results (but no raw data) available from the literature (Fig. \ref{fig:ex1_sampling}b). 
Furthermore, no information is available regarding the temperature regime for the experiment.
Transplant experiments that evaluate performance beyond the range of a species are common, and represent a plausible scenario for model integration \citep{Hargreaves2014}.
According to niche theory \citep{Holt2009}, the fundamental niche corresponds to the set of environmental conditions where the per capita intrinsic growth rate $r$ is positive.
This concept gives us a reasonable model to fit a scaling function for our sub-model (Appendix S1).
If we hypothesize that the errors from Figure \ref{fig:ex1_sampling}b \( \left(\sigma_{S} \right) \) are normally distributed, then for an observation $i$, we can interpret the probability of presence \( \left(\psi_{S,i}\right)\) as the probability that the observed growth rate \(X_{S,i}\) is positive:
\begin{equation}
	\psi_{S,i} = \int_0^\infty N \left(X_{S,i}, \sigma_{S,i} \right)
\end{equation}
where \(N\) is the Normal density function.
We can then estimate the posterior distribution for the sub-model by fitting the relationship between \(\psi_S\) and precipitation \( \left( P_S \right) \) using Bayesian beta regression \citep{Ferrari2004}:
%-----
\begin{equation}
\label{eq:ex1_thetas}
	p\left (\theta_S \mid \psi_S,P_S \right ) \propto 
	p \left(\psi_S \mid \theta_S, P_S \right)
	p \left(\theta_S \right) \\
\end{equation}
%-----

Although the two datasets were collected at considerably different scales, we now have sub-model predictions arising from a fine-scale experiment that are relevant at the scale of the metamodel (i.e., the probability of presence at a given precipitation). 
The scaling is quite simplistic; it treats the fundamental niche as the only driver of species distribution, and only considers a single dimension of the niche.
As such, it would be unwise to expect predictions from this model alone to resemble the actual distribution of the species; as a mechanistic model, it is simply too incomplete to predict distribution.
As we will show, however, the information from this sub-model, when applied as a constraint on the metamodel, can result in improved predictions that incorporate the information within each model.

We accomplish model integration by treating the posterior of \(\theta_S\) as prior information about some of the parameters of \(\theta_M\) (i.e., parameters related to precipitation), expanding Equation \ref{eq:ex1_bayes} to incorporate the new information from the sub-model:
%-----
\begin{equation}
	\label{eq:ex1_integrated}
	\overbrace{p(\theta_M \mid X_M, T_M, P_M, \theta_S, \psi_S, P_S)}^\text{integrated posterior}
	\propto
	\overbrace{p\left (\psi_S \mid \theta_S,P_S \right )}^{\substack{\text{new information} \\ \text{from sub-model}}}
	\overbrace{p \left(X_M \mid \theta_M, T_M, P_M \right) p \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
	\overbrace{p \left(\theta_S \right)}^{\substack{\text{prior for} \\ \text{sub-model}}}	
\end{equation}
%-----
As before, the metamodel \(\theta_M\) can be used to predict probability of occurrence \((\psi_I)\).
However, these predictions now reflect the presence-absence data \(X_M\) as well as the information from \(\theta_S\), including all of the data sources used to produce this sub-model.
In other words, with this evaluation, the set of parameters $\theta_M$ is most likely when it agrees with both the original data and predictions from the sub-model. 
Finally, we note the presence of marginal distributions for both models (i.e., \(p(\theta_M)\) and \(p(\theta_S)\)).
These can be informative (e.g., incorporating further prior information or the predictions of additional sub-models), semi-informative (e.g., to provide greater weight to more informative models), or uninformative.
For purposes of this example, we used these terms to reduce the precision of the second model to reflect the uncertainty of the modeling process (due to the lack of additional mechanisms in the second model, as mentioned above).

%==================
\begin{figure}[bt]
	\includegraphics{figs/ex1_precip.pdf}
	\caption{Comparison of the naive model, mechanistic sub-model, and the integrated model showing the probability of presence (\(\psi\)) as a function of precipitation.
	(a) Naive model, using only presence-absence data. Uncertainty increases dramatically when attempting to project beyond the scope of the source data.
	(b) Mechanistic model, using observations of an experiment to infer probability of presence.
	(c) Integrated model, showing predictions that are intermediate between the two sub-models and uncertainty that is reduced compared to (a).
	Uncertainty is represented as dashed lines, showing the limits of 90\% Bayesian credible intervals. The shaded region shows the calibration range for the naive model.}
	\label{fig:ex1_precip}
\end{figure}
%==================

%==================
\begin{figure}[tb]
	\includegraphics[width=5.25in]{figs/ex1_map.pdf}
	\caption{Maps showing the predicted probability of presence (\(\psi\); (a) and (c)) and the standard deviation of \(\psi\) ((b) and (d)) for the naive and integrated models.
	Historical (e.g., where presence-absence samples were available) and predicted future precipitation regimes are shown below the horizontal axes.
	Uncertainty in the naive model was extreme when predicting beyond the historical precipitation regime.
	Uncertainty in the integrated model was considerably reduced, reflecting the additional information provided by the mechanistic model.
}
	\label{fig:ex1_map}
\end{figure}
%==================

When comparing the three models (naive metamodel, mechanistic sub-model, and integrated metamodel), we observed extreme uncertainty in the first model when projecting beyond the range of the original data (Figs. \ref{fig:ex1_precip}a, \ref{fig:ex1_map}a, \ref{fig:ex1_map}b).
The sub-model was highly precise with respect to precipitation, thus providing a fairly strong constraint when producing the integrated model (Fig \ref{fig:ex1_precip}b).
The result was an integrated prediction that reflected the shapes of both models and showed considerably reduced uncertainty (Fig. \ref{fig:ex1_precip}).
At the scale of the metamodel, considering both temperature and precipitation, we observed similar results, with reduced uncertainty in the predictions over the domain not covered by the presence-absence data (Fig. \ref{fig:ex1_map}).

@@
